---
title: 'Peer Assessment 1: Activity Monitoring'
author: "juradom"
date: "February 5, 2015"
output: html_document
---
***********************************************************************
####What is mean total number of steps taken per day?

##### Process Activity Data
```{r}

# download file and unzip file
if(!file.exists("activity.csv")){
    unzip(zipfile="activity.zip")
}
# read activity data
activity.data <- read.csv("activity.csv",
                          colClasses = c("numeric", "Date","numeric")
                          )
# aggregate steps per day
activity.agg <- aggregate(steps ~ date, activity.data, 
                          sum)
```
##### Create histogram
```{r}

hist(activity.agg$steps, main = "Frequency Of Total Steps Taken Per Day", xlab = "Total Steps Per Day", ylab = "Frequency", ylim = c(0,30))
```
##### Calculate mean and median of total steps per day
```{r}
# calculate the mean total steps per day
meanSteps <- mean(activity.agg$steps)
# calculate the median total steps per day
medianSteps <- median(activity.agg$steps)
```
The mean of the total steps per day is: `r meanSteps `.
The median of the total steps per day is: `r medianSteps `.

####What is the average daily activity pattern?
##### Create Avg Steps Per Interval Plot
```{r}
# aggregate the average steps per 5 minute interval
activity.mean <- aggregate(steps ~ interval, activity.data, 
                          mean)
# create time-series plot
plot(activity.mean$interval, activity.mean$steps, main = "Average Steps Per Interval", xlab = "Interval", ylab = "Avg Steps")

```

##### Find the interval that contains the max number of steps.
```{r}
max.steps.interval <- activity.mean[activity.mean['steps']==max(activity.mean$steps)]
max.interval <- max.steps.interval[1]
```
The interval that contains the max number of steps is: `r max.interval`.