---
title: 'Peer Assessment 1: Activity Monitoring'
author: "juradom"
date: "February 5, 2015"
output: html_document
---
***********************************************************************
####What is mean total number of steps taken per day?

1) Process Activity Data
```{r}

# download file and unzip file
if(!file.exists("activity.csv")){
    unzip(zipfile="activity.zip")
}
# read activity data
activity.data <- read.csv("activity.csv",
                          colClasses = c("numeric", "Date","numeric")
                          )
# aggregate steps per day
activity.agg <- aggregate(steps ~ date, 
                          activity.data, 
                          sum)
```

2) Create histogram
```{r}

hist(activity.agg$steps, 
     main = "Frequency Of Total Steps Taken Per Day", 
     xlab = "Total Steps Per Day", 
     ylab = "Frequency", 
     ylim = c(0,30))
```

3) Calculate mean and median of total steps per day
```{r}
# calculate the mean total steps per day
meanSteps <- mean(activity.agg$steps)
# calculate the median total steps per day
medianSteps <- median(activity.agg$steps)
```
The mean of the total steps per day is: `r meanSteps `.
The median of the total steps per day is: `r medianSteps `.

***********************************************************************
####What is the average daily activity pattern?
1)  Create Avg Steps Per Interval Plot
```{r}
# aggregate the average steps per 5 minute interval
activity.mean <- aggregate(steps ~ interval, 
                           activity.data, 
                           mean)
# create time-series plot
plot(activity.mean$interval, activity.mean$steps, 
     main = "Average Steps Per Interval", 
     xlab = "Interval", 
     ylab = "Avg Steps")

```

2) Find the interval that contains the max number of steps.
```{r}
max.steps.interval <- activity.mean[activity.mean['steps']==max(activity.mean$steps)]
max.interval <- max.steps.interval[1]
```
The interval that contains the max number of steps is: `r max.interval`.

***********************************************************************
####Imputing missing values

1) Calculate the total number of missing values
```{r}
library(plyr)
# find rows with NAs
missing.activities.counts <- nrow(count(which(is.na(activity.data))))
```
The total number of missing values in the dataset is: `r missing.activities.counts `.

2) Populates NAs with average steps for that 5 min interval
```{r}
library(dplyr)
imput.activity.data <- merge(activity.data, activity.mean, by.x = "interval", by.y ="interval")
complete.activity.data <- imput.activity.data %>%
    group_by(interval) %>%
    mutate(steps.x = ifelse(is.na(steps.x),
                          steps.y,
                          steps.x))

```
3) Create histogram
```{r}
# aggregate steps per day
complete.activity.agg <- aggregate(steps.x ~ date, 
                          complete.activity.data, 
                          sum)

hist(complete.activity.agg$steps.x, 
     main = "Frequency Of Total Steps Taken Per Day", 
     xlab = "Total Steps Per Day", 
     ylab = "Frequency")
```

4) Calculate mean and median of total steps per day
```{r}
# calculate the mean total steps per day
meanCompleteSteps <- mean(complete.activity.agg$steps)
# calculate the median total steps per day
medianCompleteSteps <- median(complete.activity.agg$steps)
```
The mean of the total steps per day is: `r meanCompleteSteps `.
The median of the total steps per day is: `r medianCompleteSteps `.

***********************************************************************
####Are there differences in activity patterns between weekdays and weekends?

1) Create time-series plot
```{r}
library(data.table)
library(lattice)

# Create weekday/weekend factor variable
weekday.activity <- data.table(complete.activity.data)
weekday.activity <- weekday.activity[,weekpart := ifelse(weekdays(date)  %in% c("Saturday", "Sunday"),"weekend","weekday")]

# Create time-series plot
# aggregate the average steps per 5 minute interval
weekday.activity.mean <- aggregate(steps.x ~ interval + weekpart, 
                           weekday.activity, 
                           mean)
# create time-series plot
xyplot(steps.x ~ interval | weekpart , 
       data=weekday.activity.mean,
       layout=c(1,2),
       ylab="Number Of Steps",
       xlab="Interval",
       type="l")

```


